Causal graph analysis of COVID-19 observational data in German districts reveals effects of determining factors on reported case numbers

### 2021-04-19 manuscript/v6_medrxiv_revision_2/discussion.Rmd
update discussion social determinants

@@

Furthermore, our analyses strongly support the effectiveness of non-pharmaceutical interventions. To a lesser extent, the adverse effects of some socio-demographic factors might help to identify areas that are at higher risk of local COVID-19 outbreaks.

>>>>

A recent review on COVID-19 based on evidence from the US and UK concludes that low socio-economic status groups are being hit harder by the pandemic [@Wachtler2020Socioeconomic]. Albeit specific pathways remain unclear, many studies found associations with poverty or its correlates such as poor and potentially overcrowded housing conditions. For Germany, a higher case fatality of COVID-19 cases in districts with higher socio-economic deprivation has also been reported just recently, which was especially pronounced in the second wave of the pandemic [@rki2021sozialeunterschiede]. Similarly, our analysis identified a decreasing effect on COVID-19 case numbers within districts with a higher socio-economic status during the first wave.

@@

Furthermore, our analyses strongly support the effectiveness of non-pharmaceutical interventions. To a lesser extent, the adverse effects of some socio-demographic factors might help to identify areas that are at higher risk of local COVID-19 outbreaks and more severe outcomes of infection cases.
